\section{Applied uses}

In health technology assessment, an analyst can estimate whether uncertainty
about treatment effects, adverse events, quality of life, resource use, or
costs is important enough to justify another study. EVSI and ENBS can compare
candidate sample sizes or designs; implementation analysis can separately ask
whether improving uptake would create more value than collecting more
evidence. Heterogeneity can identify circumstances in which a treatment is
worthwhile for one observable group but not another, while equity weighting
can examine the trade-off between total health and its distribution.

Genomic testing provides one concrete example of why the wider decision
context matters. Process-level microcosting identifies which resources drive
the cost of testing~\cite{mordaunt2024microcosting}; observed uptake after
public funding shows that availability does not imply uniform
implementation~\cite{mordaunt2023uptake}; and diffusion-based budget-impact
analysis represents how adoption and expenditure develop over
time~\cite{mordaunt2026budget}. These inputs address different parts of a
decision model and should not be collapsed into parameter uncertainty alone.

In public policy, the same framework can compare a pilot, wider data
collection, immediate rollout, targeted implementation, or waiting. Relevant
uncertainty may concern programme effects, behavioural response, delivery
capacity, costs falling on different groups, or whether evidence from one
jurisdiction transports to another. The result is meaningful only when the
model represents the policy alternatives, institutional constraints, affected
population, and chosen public-value objective.

In marketing and business analytics, a decision might concern product launch,
pricing, campaign allocation, customer targeting, capacity, or market entry.
Information can come from an experiment, survey, forecast improvement, or
additional observation of demand. VOI complements predictive accuracy by
valuing whether better predictions would actually change an action and improve
expected business value after experimentation cost and delay. It does not
convert a predictive association into a causal effect or remove the need to
model competitive response and implementation.

Environmental, energy, engineering, manufacturing, finance, and portfolio
applications can use the same logic for monitoring, mitigation, inspection,
maintenance, reliability, quality control, forecasting, or irreversible
investment. These applications have not yet been empirically validated by this
project. They require domain-appropriate causal models, constraints, outcomes,
and value functions; the package supplies decision analysis, not domain facts.

Developing features also address evidence-governance questions: whether
validation data could change deployment, whether source evidence transports to
a target population, whether linkage or measurement improvements justify
their cost and privacy burden, and whether aging evidence warrants
reassessment. These formulations can make the value judgments and
counterfactual choices explicit, but they require further methodological and
applied validation.

Version 1.0.0, released under Apache-2.0, is available from
\url{https://github.com/edithatogo/voiage/releases/tag/v1.0.0}~\cite{voiage2026}.
Its repository
snapshot is archived by Software Heritage as
\url{swh:1:snp:767efde24c97d9f6d730764c1b3bc1a91ba20c32}. Intended uses do not
constitute evidence of adoption or policy impact, and no example replaces
validation with relevant domain data and decision-makers.
