\section{Economic interpretation and applied analysis}

\subsection{From uncertainty to opportunity loss}

The information measures have a common economic interpretation. A decision
made with current information can select a strategy that is not best for the
state of the world that is eventually realized. The resulting opportunity loss
is the difference between the value of the best strategy for that state and
the value of the strategy selected using current evidence. EVPI averages the
avoidable part of this loss if all relevant uncertainty could be resolved.
EVPPI asks the same question for a subset of uncertainty, while EVSI values the
information expected from a particular study.

These measures answer different policy questions. EVPI places an upper bound
on the value of eliminating the uncertainty represented in the model. EVPPI
helps prioritize which uncertainties merit investigation. EVSI compares
specific study designs under a model of how data would update beliefs and
change action. ENBS then subtracts the cost of obtaining that information. A
positive ENBS supports the study only under the stated population, timing,
implementation, and cost assumptions; it is not a general statement that more
research is socially worthwhile.

\subsection{Research, implementation, and delay}

The value of information is not necessarily the value that will be realized.
If a preferred strategy will reach only part of the eligible population, or
will be adopted slowly or followed incompletely, the gain from improving the
decision can be smaller than the gain predicted under full implementation.
Conversely, implementation activity may be more valuable than another study
when evidence is already sufficient but delivery is weak. The two investments
should therefore be compared explicitly rather than combining evidence
generation and implementation into one unexplained adjustment.

Population scaling is similarly consequential. The number of future decisions
affected by information depends on incidence or eligibility, the lifetime of
the decision, discounting, adoption, displacement by later technologies, and
the delay before results become available. A large per-person EVSI can have
little population value for a rare or short-lived decision. A small per-person
value can justify research when it affects many decisions over a long period.
The calculation in Equation~(4) deliberately exposes a simple constant-
population assumption; applications with changing populations or delayed
availability should supply an appropriate decision-opportunity model.

\begin{table}[!ht]
\centering
\footnotesize
\begin{tabular}{p{0.20\linewidth}p{0.34\linewidth}p{0.36\linewidth}}
\toprule
Source of avoidable loss & Decision question & Relevant analysis \\
\midrule
Information & What would perfect, partial, or sample information change? & EVPI, EVPPI, EVSI, ENBS \\
Model & Is uncertainty in structure, validation, or computation decision-relevant? & Structural VOI, model validation, computational refinement \\
Population & Would subgroups, preferences, perspective, or equity change action? & Heterogeneity, distributional equity, preference, perspective \\
Delivery & Will adoption and execution realize modeled value? & Implementation adjustment and strategy comparison \\
Evidence system & Is evidence usable in the target decision? & Data quality, transportability, synthesis, monitoring, obsolescence \\
Timing & Should evidence collection or commitment adapt over time? & Sequential and adaptive studies, dynamic real options, portfolios \\
\bottomrule
\end{tabular}
\caption{Sources of avoidable decision loss represented in version 1. Inclusion
means that the question can be represented, not that every method has the same
level of validation.}\label{tab:problems}
\end{table}
\FloatBarrier{}

\subsection{Heterogeneity, equity, and perspective}

Population averages can hide different preferred actions. Subgroup analysis
asks whether decisions informed by observable group membership improve
expected value. Its usefulness depends on whether the groups can be identified
at the point of decision, whether subgroup effects are credible, and whether
the sample represents the target population. Uncertain or data-driven subgroup
selection can otherwise exaggerate apparent gains.

Equity analysis changes the objective rather than merely adding another
covariate. A distributional analysis may assign different social weights to
health or welfare gains, or examine the opportunity cost of reducing
inequality~\cite{asaria2016distributional}. The preferred strategy can then
differ from the strategy that maximizes unweighted total net benefit.
\texttt{voiage} keeps subgroup weights and the analytical perspective visible
because neither is a neutral technical setting. A payer, health-system,
patient, societal, firm, or public-policy perspective can include different
costs, outcomes, constraints, and time horizons.

\subsection{Model and evidence uncertainty}

Parameter VOI is conditional on the model being an adequate account of the
decision. Structural VOI instead asks whether choosing among plausible model
forms could change action. Transportability asks whether evidence from its
source population is informative for the target population. Data-quality and
validation analyses ask whether improving measurement, linkage, calibration,
or external validation could reduce decision loss. These quantities require
substantive alternatives: arbitrary model variants or unsupported penalties do
not become meaningful merely because a numerical value can be calculated.

\subsection{Worked decision example}

For two strategies with net-benefit rows $(0,100)$ and $(0,-100)$, the sample
EVPI is 50 value units per decision. Perfect information would identify the
better strategy in each state, whereas current evidence makes the strategies
equally valuable on average. Suppose a feasible study has an EVSI of 12.5 per
future decision. This is one quarter of EVPI: the study is expected to resolve
some, but not all, consequential uncertainty.

If the study affects 1000 decisions each year for 10 years, using a 3\% annual
discount rate gives 8530.2028 discounted decision opportunities. Before
research cost, the expected population value is therefore 106627.5355. With a
total study cost of 5, population ENBS is 106622.5355 value units. The example
is intentionally simple. If study results arrive after two years, the early
decisions receive no benefit from that information; if uptake is 60\%, the
realized value may also be lower. Those assumptions can matter more to the
investment decision than additional decimal precision in the EVSI estimate.

\subsection{Analytical workflow and checks}

An applied analysis begins by defining the strategies, target population,
perspective, outcome-value scale, decision lifetime, and uncertain states. The
analyst then identifies the decision that information could change and the
counterfactual intervention: perfect information, a particular study, better
implementation, improved measurement, or delayed commitment. Only after these
choices are explicit should the numerical estimator and population scaling be
selected.

\texttt{voiage} retains labels for samples, strategies, uncertain inputs,
subgroups, perspectives, populations, and study designs using labelled data
conventions~\cite{hoyer2017xarray}. Results can retain the preferred strategy,
method, random seed, warnings, scaling assumptions, and data sources. Checks
cover properties such as invariance to a common shift in net benefit,
non-negativity where theory requires it, repeatability, difficult statistical
models, identical strategies, and malformed inputs. Approximate EVPPI or EVSI
estimates can still be distorted by Monte Carlo error, poor model fit, or
extrapolation. No software check can establish that the decision model,
evidence, perspective, or social-value judgments are appropriate.

Selected established numerical components are implemented in Rust and made
available through the supported language packages. This implementation detail
is secondary to the analytical contract: equivalent inputs should represent
the same decision and return numerically consistent results. Python currently
supports the widest workflow; the R and Julia packages provide a narrower
subset.
