\section{Statement of need}

VOI analysis can inform whether additional research is worthwhile. Practical
implementations may be split across domain-specific packages, languages, and
model formats, which can complicate comparison and obscure which uncertainty
is being valued. A parameter-level
EVPPI does not answer whether the model structure is wrong; an EVSI does not by
itself account for incomplete implementation; and population-average VOI can
hide subgroup-specific decisions or equity trade-offs. Structural uncertainty,
heterogeneity, and implementation are established methodological concerns, not
software edge cases~\cite{jackson2011structural,espinoza2014heterogeneity,fenwick2008implementation}.

\texttt{voiage} addresses five connected decision questions:
\begin{enumerate}
  \item \emph{parameter and sample information}: which uncertain inputs or
  proposed studies could change the preferred strategy;
  \item \emph{model and evidence adequacy}: whether structural assumptions,
  validation data, measurement quality, or transportability warrant further
  investment;
  \item \emph{population and social value}: whether subgroup heterogeneity,
  preferences, equity weights, or analytical perspective alter the decision;
  \item \emph{delivery}: whether uptake, adherence, coverage, delay, or
  implementation uncertainty erodes expected value; and
  \item \emph{timing and control}: whether sequential learning, adaptive
  sampling, waiting, irreversibility, or portfolio constraints change when to
  act.
\end{enumerate}

This taxonomy is domain-neutral, but its inputs retain domain meaning. Health
analyses can use net monetary or health benefit, willingness-to-pay thresholds,
eligible populations, and trial costs. Environmental analyses can compare
monitoring or mitigation strategies under ecological and monetary
consequences. Engineering and manufacturing analyses can value reliability,
failure, inspection, or quality information. Financial and energy analyses can
value delayed commitment, changing evidence, and irreversible investment. The
common requirement is an explicit mapping from uncertain states and strategies
to comparable decision value; the package does not pretend that the mapping is
scientifically interchangeable across domains.

Let $NB_{sd}$ denote the net benefit for probabilistic sample $s$ and strategy
$d$. For $S$ samples and $D$ strategies, the sample EVPI estimator is
\begin{equation}
  \widehat{\mathrm{EVPI}} =
  \frac{1}{S}\sum_{s=1}^{S}\max_{d} NB_{sd}
  - \max_{d}\frac{1}{S}\sum_{s=1}^{S} NB_{sd}.
\end{equation}
This estimates the expected gain from perfect information, equivalently the
expected reduction in decision loss relative to acting with current
information. Under representative independent and identically distributed
draws with finite expectations, the plug-in estimator is consistent. In finite
samples it is generally downward-biased because the maximum of estimated
strategy means is upward-biased; its Monte Carlo uncertainty should therefore
be assessed. EVPPI restricts perfect information to a parameter subset $\phi$
and integrates over the complementary uncertainty $\theta_{-\phi}$,
\begin{equation}
  \mathrm{EVPPI}(\phi) =
  \mathbb{E}_{\phi}\!\left[\max_d
  \mathbb{E}_{\theta_{-\phi}\mid\phi}
  \{NB_d(\phi,\theta_{-\phi})\}\right]
  - \max_d \mathbb{E}_{\theta}\{NB_d(\theta)\},
\end{equation}
which is commonly estimated using regression from a probabilistic sensitivity
analysis sample~\cite{strong2014evppi}. EVSI instead values a proposed finite
study with fixed design and data $Y$ by averaging the decision made after
observing its data,
\begin{equation}
  \mathrm{EVSI}(Y) =
  \mathbb{E}_{Y\sim p(Y)}\!\left[\max_d
  \mathbb{E}_{\theta\mid Y}\{NB_d(\theta)\}\right]
  - \max_d \mathbb{E}_{\theta}\{NB_d(\theta)\}.
\end{equation}
Regression-based reuse of existing simulation output can reduce the cost of
this calculation~\cite{strong2015evsi}. The supplied nested Monte Carlo path
targets this preposterior estimand for a conjugate normal study model. Other
exposed regression, moment, and efficient approximations make additional
modeling or heuristic assumptions; numerical bounding can conceal Monte Carlo
or model-fit error and is not evidence of calibration.

For per-decision, per-period EVSI $v$, recurring population $N$, horizon $T$,
annual discount rate $r$, and total research cost $C$, \texttt{voiage} defines
\begin{equation}
  \mathrm{ENBS}_{\mathrm{population}} =
  vN\left(\frac{1-{(1+r)}^{-T}}{r}\right)-C,
\end{equation}
with the annuity factor replaced by $T$ when $r=0$. This scaling assumes
constant recurring decision opportunities, end-of-period discounting, no
population change, and no implementation delay. It must not be applied to
values already aggregated over population or time. ENBS remains signed;
$\max(0,\mathrm{ENBS})$ is the separate value of the option to commission the
study.

Implementation choices for these estimands can differ. Examples include the
orientation of samples and strategies, handling of non-finite values,
tie-breaking rules, and approximation methods. The \texttt{voiage} numerical
contract therefore records dimensions, finite-value checks, method metadata,
warnings, seeds, and serialized result envelopes.

The same principle extends beyond classical estimands. For example,
equity-weighted subgroup aggregation compares a single population-wide
decision with subgroup-specific optima and an explicitly weighted
social-welfare decision~\cite{asaria2016distributional}. Its interpretation is
population-relevant only when subgroup weights represent the target
population; sample-frequency weights yield a sample-weighted contrast. The
basic implementation adjustment is a deterministic multiplier or scenario
calculation for uptake, adherence, coverage, delay, and an uncertainty penalty.
A common positive multiplier does not change strategy ranking and is not a
probabilistic value-of-implementation estimand. Sequential and adaptive
analyses compare study designs and stopping rules through their expected net
benefit of sampling, consistent with the decision-theoretic role of EVSI in
trial design~\cite{flight2022sequential}.
