\section{Decision problem and related work}

VOI is established in statistical decision analysis and health-economic
evaluation~\cite{claxton1999irrelevance,briggs2006decision}. R tooling includes
the Bayesian cost-effectiveness analysis (BCEA)
ecosystem~\cite{baio2017bcea}, while regression estimators have made EVPPI and
EVSI more practical~\cite{strong2014evppi,strong2015evsi}. Methods also address
structural uncertainty, heterogeneity, distributional cost-effectiveness,
implementation, transportability, adaptive study design, and the value of
delaying a reversible decision~\cite{jackson2011structural,espinoza2014heterogeneity,asaria2016distributional,fenwick2008implementation,bareinboim2016datafusion,flight2022sequential,palmer2000options}.
These methods answer related but different questions. Reimplementing their
names does not constitute a methodological contribution, and
\texttt{voiage} does not claim statistical superiority over specialist
implementations.

Practical analyses may be split across domain-specific packages, languages,
and model formats. This can obscure which source of uncertainty is being
valued. Parameter-level EVPPI does not show whether another model structure
would change the decision. EVSI does not by itself account for incomplete
implementation. Population-average VOI can hide subgroup decisions or equity
trade-offs. Evidence from one population may also have limited value in
another when effects, costs, preferences, or delivery differ.

\texttt{voiage} groups these concerns into five decision questions:
\begin{enumerate}
  \item \emph{parameter and sample information}: which uncertain inputs or
  proposed studies could change the preferred strategy;
  \item \emph{model and evidence adequacy}: whether structural assumptions,
  validation data, measurement quality, or transportability warrant further
  investigation;
  \item \emph{population and social value}: whether heterogeneity, preferences,
  equity weights, or analytical perspective alter the decision;
  \item \emph{delivery}: whether uptake, adherence, coverage, or delay changes
  the value likely to be realised; and
  \item \emph{timing and control}: whether sequential learning, waiting,
  irreversibility, or portfolio constraints change when to act.
\end{enumerate}

These concerns can interact. Research may have little realised value if
implementation is weak. A strategy that maximises average net benefit may not
maximise an equity-weighted objective. Immediate adoption may remove the value
of waiting when a decision is costly to reverse. The quantities are not
automatically commensurable: each depends on a value function, target
population, time horizon, counterfactual action, and evidence model.

The framework is domain-neutral, but its inputs retain domain meaning. Health
analyses can use net monetary or health benefit, willingness-to-pay thresholds,
eligible populations, and study costs. Other applications can use ecological,
engineering, business, or public-value outcomes when uncertain states and
strategies map to a comparable decision-value scale. The shared computational
form does not make those substantive mappings interchangeable.

\begin{table}[!ht]
\centering
\footnotesize
\begin{tabular}{p{0.20\linewidth}p{0.34\linewidth}p{0.36\linewidth}}
\toprule
Source of avoidable loss & Decision question & Relevant analysis \\
\midrule
Information & What would perfect, partial, or sample information change? & EVPI, EVPPI, EVSI, ENBS \\
Model & Could structure, validation, or computation change the decision? & Structural VOI, validation, computational refinement \\
Population & Would subgroups, preferences, perspective, or equity change action? & Heterogeneity, distributional equity, preference, perspective \\
Delivery & How much modelled value is likely to be realised? & Implementation adjustment and strategy comparison \\
Evidence system & Is evidence suitable for the target decision? & Data quality, transportability, synthesis, monitoring, obsolescence \\
Timing & Could evidence collection or commitment adapt over time? & Sequential studies, real options, portfolios \\
\bottomrule
\end{tabular}
\caption{Sources of avoidable decision loss represented in version 1. The
table describes analytical scope; individual methods retain established,
developing, or experimental status.}\label{tab:problems}
\end{table}
\FloatBarrier{}
