---
title: "Introduction to voiageR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to voiageR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# voiageR: Value of Information Analysis in R

The `voiageR` package provides an R interface to the `voiage` Python library for Value of Information (VOI) analysis. This vignette demonstrates how to use the package to perform common VOI analyses.

## Installation

To use `voiageR`, you first need to install the Python `voiage` package:

```bash
pip install voiage
```

Then you can install `voiageR` from GitHub:

```r
# install.packages("devtools")
devtools::install_github("edithatogo/voiage", subdir = "r-package/voiageR")
```

## Basic Usage

### Calculating EVPI

The Expected Value of Perfect Information (EVPI) represents the maximum amount a decision-maker should be willing to pay to eliminate all uncertainty in a decision model.

```{r eval=FALSE}
library(voiageR)

# Create sample net benefit data
# 1000 PSA samples, 2 strategies
net_benefits <- matrix(rnorm(2000), nrow = 1000, ncol = 2)

# Calculate EVPI
evpi_value <- evpi(net_benefits)
print(evpi_value)
```

You can also scale the EVPI to a population:

```{r eval=FALSE}
# Calculate population-level EVPI
evpi_pop <- evpi(
  net_benefits = net_benefits,
  population = 100000,
  time_horizon = 10,
  discount_rate = 0.03
)
print(evpi_pop)
```

### Calculating EVPPI

The Expected Value of Partial Perfect Information (EVPPI) quantifies the value of learning the true value of a specific subset of model parameters.

```{r eval=FALSE}
# Create parameter samples
param_samples <- list(
  param1 = rnorm(1000),
  param2 = rnorm(1000)
)

# Calculate EVPPI
evppi_value <- evppi(net_benefits, param_samples)
print(evppi_value)
```

### Calculating EVSI

The Expected Value of Sample Information (EVSI) estimates the value of conducting a specific study to reduce uncertainty.

```{r eval=FALSE}
model_func <- function(params) {
  cbind(
    Treatment = 50000 * params$mean_treatment - 3000,
    Control = 50000 * params$mean_control
  )
}

set.seed(17)
draws <- 1000L
prior_samples <- list(
  mean_treatment = rnorm(draws, mean = 0.06, sd = 0.03),
  mean_control = rnorm(draws, mean = 0.00, sd = 0.02),
  sd_outcome = rep(1.0, draws)
)

trial_design <- list(
  arms = list(
    list(name = "Treatment", sample_size = 100L),
    list(name = "Control", sample_size = 100L)
  )
)

evsi_value <- evsi(
  model_func = model_func,
  prior_samples = prior_samples,
  trial_design = trial_design,
  method = "two_loop",
  n_outer_loops = 100L,
  n_inner_loops = 1000L,
  seed = 20260724L
)
print(evsi_value)
```

The built-in two-loop model lowercases each arm name and replaces spaces with
underscores to obtain the required `mean_<normalised arm>` parameter name. It
also requires a finite, strictly positive `sd_outcome` fixed across all prior
draws.

Custom trial-simulation and posterior-sampling callbacks remain Python-only.
The R facade does not convert or execute those callbacks and does not claim
custom-model parity.

## Advanced Features

### Using Different Python Environments

If you have installed `voiage` in a specific Python environment, you can specify which environment to use:

```{r eval=FALSE}
# Use a virtual environment
set_voiage_env("myenv", type = "virtualenv")

# Use a conda environment
set_voiage_env("myenv", type = "conda")
```

### Checking Package Availability

You can check if the `voiage` Python package is available:

```{r eval=FALSE}
is_available <- is_voiage_available()
print(is_available)
```

## Conclusion

The `voiageR` package provides a seamless interface between R and the powerful `voiage` Python library for Value of Information analysis. This allows R users to leverage advanced VOI methods while working in their preferred environment.
