---
title: "Getting Started with voiageR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with voiageR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  purl = FALSE,
  echo = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

`voiageR` is the thin R interface to the `voiage` Python library. It exposes
the stable EVPI, EVPPI, and EVSI wrappers while keeping the package surface
small, predictable, and suitable for scripted use.

The examples below are intentionally non-interactive. They show the package
shape, the Python environment setup, and the canonical data contracts without
requiring the vignette build to depend on a live Python session.

# Set up the Python environment

```{r setup-env, purl=FALSE}
library(voiageR)

set_voiage_env("voiage", type = "virtualenv")
init_voiage()
```

If you manage Python through Conda, use:

```{r setup-conda, purl=FALSE}
set_voiage_env("voiage", type = "conda")
```

# EVPI

```{r evpi, purl=FALSE}
net_benefits <- matrix(
  c(10.0, 12.0, 11.0, 9.0, 13.0, 14.0),
  nrow = 3,
  byrow = TRUE,
  dimnames = list(NULL, c("Strategy A", "Strategy B"))
)

evpi(net_benefits)
```

# EVPPI

```{r evppi, purl=FALSE}
parameter_samples <- list(
  effect = c(0.10, 0.12, 0.11),
  cost = c(1000, 980, 1025)
)

evppi(net_benefits, parameter_samples, parameters_of_interest = "effect")
```

# EVSI

```{r evsi, purl=FALSE}
model_func <- function(params) {
  cbind(
    Treatment = 50000 * params$mean_treatment - 3000,
    Control = 50000 * params$mean_control
  )
}

set.seed(17)
draws <- 1000L
prior_samples <- list(
  mean_treatment = rnorm(draws, mean = 0.06, sd = 0.03),
  mean_control = rnorm(draws, mean = 0.00, sd = 0.02),
  sd_outcome = rep(1.0, draws)
)

trial_design <- list(
  arms = list(
    list(name = "Treatment", sample_size = 100L),
    list(name = "Control", sample_size = 100L)
  )
)

evsi(
  model_func,
  prior_samples,
  trial_design,
  method = "two_loop",
  n_outer_loops = 100L,
  n_inner_loops = 1000L,
  seed = 20260724L
)
```

The built-in model expects one `mean_<normalised arm>` parameter for every
trial arm. It lowercases each arm name and replaces spaces with underscores.
`sd_outcome` is the known outcome standard deviation and is fixed across the
prior draws.

Custom trial-simulation and posterior-sampling callbacks remain Python-only.
The R facade does not convert or execute those callbacks and does not claim
custom-model parity.

# Current wrapper surface

The current R package exports the core analysis wrappers only:

- `evpi()` for perfect-information value
- `evppi()` for parameter-specific information value
- `evsi()` for the built-in two-loop model and compatibility estimators

Composite metrics such as ENBS are currently part of the Python surface and
can be composed from the results above when needed.

# Build outputs

The package ships a deterministic PDF manual build helper at
`tools/build-manual.R`. It is designed for non-interactive use from the package
root and produces a stable manual artifact from the package Rd files for
release packaging and local verification.
