\documentclass[11pt]{article}

% Prefer packages supported by arXiv's TeX Live and LaTeXML environments.
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\emergencystretch}{2em}
\hypersetup{pdftitle={voiage: Value of Information for Research, Implementation, and Decision Making},
  pdfauthor={Dylan A. Mordaunt},
  pdfsubject={Decision analysis under parameter, structural, population, implementation, and timing uncertainty},
  pdfkeywords={value of information, decision analysis, structural uncertainty, distributional analysis, implementation, sequential decisions}
}

\title{\texttt{voiage}: Value of Information for Research, Implementation,
and Decision Making}
\author{Dylan A. Mordaunt\textsuperscript{1,2,3}\\
  \footnotesize\textsuperscript{1}Faculty of Health, Education and Psychology, Victoria University of Wellington\\
  \footnotesize\textsuperscript{2}College of Medicine and Public Health, Flinders University\\
  \footnotesize\textsuperscript{3}Centre for Health Policy, The University of Melbourne}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Value of Information (VOI) analysis helps decision-makers judge whether
reducing uncertainty is worth the cost. Many analyses focus on uncertain
numerical inputs but do not show whether the answer could also change because
of the model used, differences between population groups, the viewpoint
adopted, poor implementation, weak evidence, or the timing of a decision.
\texttt{voiage} provides a common way to describe and analyse these connected
questions. Its established features calculate the expected value of perfect,
partial, and sample information, the net value of further research, dominance,
cost-effectiveness acceptability frontiers, and differences between groups.
Newer features examine questions such as model choice, equity, implementation,
evidence quality, adaptation, and delayed or irreversible action. The software
keeps the available choices, relevant groups and populations, assumptions,
warnings, and data sources visible with the result. It does not replace the
scientific judgment needed to build a suitable model or value outcomes. This
paper explains the economic questions addressed by version 1.0.0, the
interpretation of its information measures, and where further validation is
still needed.
\end{abstract}

\input{sections/summary}
\input{sections/statement_of_need}
\input{sections/state_of_the_field}
\input{sections/software_design}
\input{sections/research_impact}
\input{sections/limitations}
\input{sections/acknowledgements}
\input{sections/glossary}
\input{sections/abbreviations}

\bibliographystyle{unsrt}
\begingroup
\small
\bibliography{references}
\endgroup

\end{document}
